---
title: ""
author: "Me"
data: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include=FALSE}
setwd("G:/My Drive/Work/TB Projects/ReMoTe")
source("./R/utils.R")
library(ggplot2)
library(reshape)
library(reshape2)
library(RColorBrewer)
library(ggpubr)
library(gridExtra)
library(grid)
library(scales)
library(dplyr)
library(data.table)
library(xlsx)
library(kableExtra)
```

## Background

We initialize a population of size one million using the ARI, and age structure that mimics South Africa. We then generate new case of disease over a one year period and explore the role of reinfection in generating new cases of disease. We simulate populations with prevalence ranging from 100/100k to 1000/100k.

We consider two scenarios: 

1. Homogenous risk of TB infection in the population
2. Hetergenous risk: 20% of the population has an ARI that is 6 times greater than the low risk population. 

The ARI is challenging to estimate. We use two approaches:

1. Dodd et al demonstrate a relationship between TB prevalence and the ARI. We leverage this to determine the overall ARI.
2. We assume that the duration of disease is two years, as Styblo, and therefore prevalence must be twice incidence. In this case, we do parameter sweeps to determine the ARI that meets this condition for each prevalence scenario. 

We also use the correction in these calculations that true prevalence should be 1.66 times reported to account for smear negative TB.

We keep track of repeat infection events and, following Ackley et al and Reichler, we allow there to initially be a protective effect of reinfection, but that this gradually wanes with an increasing number of infections. Following is a depiction of this assumption:

```{r}
max.inf <- 100
TB.prog=c(0.04,0.02,0.01)
TB.prog.sim <- c(TB.prog[1],seq(TB.prog[2],TB.prog[2]+TB.prog[3]*(max.inf-1),TB.prog[3]))
TB.prog.sim[TB.prog.sim>1] <- 1

plot.dat <- data.frame(num.infections=seq(1,max.inf+1),prog.rate=TB.prog.sim)

ggplot(plot.dat,aes(x=num.infections,y=prog.rate))+
  geom_line()+
  labs(x="Number of infections",y="Probability of pregression to disease")

```

## Simulation scenarios
We run the simulation with four total scenarios: two ARI scenarios and two mixing scenarios.

ARI scenarios:

-Dodd et al ARIs (based on TB prevalence)

-ARIs derived assuming two year duration

Mixing scenarios:

-All population subject to the same risk of infection

-20% of the population is in a high risk category with 6 times the risk of infection as the remaining 80% of the population.

Following are these parameter assumptions:

```{r, echo = FALSE}
# Dodd et al ARI, based on TB prevalence
pars.mix <- read.xlsx("G:/My Drive/Work/TB Projects/ReMoTe local/Code/ParameterInputs/New Baseline parameters for model_fixed HR_9.16.21 standard.xlsx",sheetName = "Mixing and Ackley")
# get overall value
ari.overall.mix <- round(pars.mix[1:10,2]*pars.mix[1:10,4]+pars.mix[1:10,3]*pars.mix[1:10,5],2)

pars.nomix <- read.xlsx("G:/My Drive/Work/TB Projects/ReMoTe local/Code/ParameterInputs/New Baseline parameters for model_fixed HR_9.16.21 standard.xlsx",sheetName = "No Mixing, Ackley")
# 10/12/21: do not use this-just use overall ARI for mixing scenarios

# par sweep parameters
pars.sweep <- read.csv("G:/My Drive/Work/TB Projects/ReMoTe/Parameters/parSweepAll_2yr.csv")

pars.print <- data.frame(cbind(pars.mix[1:10,1:3],ari.overall.mix,ari.overall.mix,
                               pars.sweep[,c(8,9,7,5)]))

pars.print %>% kbl(col.names=c("prevalence","Dodd: ARI low risk","Dodd: ARI high risk","Dodd: overall ARI, mixing","Dodd: no mix ARI",
                               "Sweep: ARI low risk","Sweep: ARI high risk","Sweep: overall ARI, mixing","Sweep: no mix ARI")) %>% kable_styling()

pop.age=c(99000, 99000,92000,83000,85000,90000, 91000,79000,63000,54000,44000,37000,30000, 23000,15000,10000,5000,1000)
median.age=c(2,7,12,17,22,27,32,37,42,47,52,57,62,67,72,77,82,87)
prog.old.inf=0.00075
```

## Initializing the population {.tabset}

We first determine the number of infections that individuals have at the start of the simulation. We show this for all four scenarios.

```{r, echo=F}
# simulate the baseline population
sim.dodd.mix <- sim.dodd.nomix <- sim.sweep.mix <- sim.sweep.nomix <- list()
for(i in 1:length(pars.mix[,1])){
  
  if(is.na(pars.mix[i,1]) || is.na(pars.nomix[i,1])){break}
  pars.tmp.mix <- pars.mix[i,]
  sim.dodd.mix[[i]] <- init.popn.mix(ari.high=pars.tmp.mix$ARI.high.risk/100,
                                ari.low=pars.tmp.mix$ARI.low.risk/100,
                                low.prev=pars.tmp.mix$percent.low.risk)

  pars.tmp.nomix <- ari.overall.mix[i]
  sim.dodd.nomix[[i]] <- init.popn.nomix(ari=pars.tmp.nomix/100)

  sim.sweep.mix[[i]] <- init.popn.mix(ari.high=pars.sweep$high.ARI..mix[i]/100,
                                ari.low=pars.sweep$low.ARI..mix[i]/100,
                                low.prev=pars.tmp.mix$percent.low.risk)

  sim.sweep.nomix[[i]] <- init.popn.nomix(ari=pars.sweep$overall.ARI..no.mix[i]/100)
}  
```

```{r, echo=FALSE}
# first show overall number of infections by prevalence
# need matrix with prevalence and number of infections, then have this by age group
num.prev <- length(unique(pars.mix[,1])); num.age <-length(median.age)
# data frame for plotting with age,prev,total infections
num.inf.age.mix.dodd <- num.inf.age.nomix.dodd <-
  num.inf.age.mix.sweep <- num.inf.age.nomix.sweep <- data.frame(prevalence=rep(pars.mix[1:10,1],each=num.age+1),
                             age.group=rep(c(0,median.age),num.prev),
                             total.infections=rep(NA,num.prev*(num.age+1)))

# data frame with prevalence,number of infections (categorical), count with that number of infections/prev
num.inf.mix.dodd <- num.inf.mix.hi.dodd <- num.inf.mix.lo.dodd <- num.inf.nomix.dodd <- 
  num.inf.mix.sweep <- num.inf.mix.hi.sweep <- num.inf.mix.lo.sweep <- num.inf.nomix.sweep <- data.frame(prevalence=rep(pars.mix[1:10,1],each=10),
                             num.inf=rep(c(as.character(c(0:8)),"9+"),num.prev),
                             count=rep(NA,num.prev*10))

# data frame for high/low and number of infections
num.inf.hi.low.dodd <- num.inf.hi.low.sweep <- data.frame(prevalence=rep(pars.mix[1:10,1],each=2),
                             risk.cat=rep(c("High","Low"),10),
                             total.infections=rep(NA,num.prev*2))

# data frame prev of LTBI for each prevalenceTB x age combo
age.prev.LTBI.mix.dodd <- age.prev.LTBI.nomix.dodd <- 
  age.prev.LTBI.mix.sweep <- age.prev.LTBI.nomix.sweep <- matrix(0,nr=num.age,nc=10)

for(i in 1:10){
  
  # total number of infected individuals in each age group by prevalence
  #########################################################################
  ## Dodd assumption
  tmp.inf <- sim.dodd.mix[[i]]$all.pop
  tmp.inf <- tmp.inf[,2:ncol(tmp.inf)]
  ### total.infections: (total  infections, age specific number of infections)
  num.inf.age.mix.dodd$total.infections[((i-1)*(num.age+1)+1):(i*(num.age+1))] <- round(c(sum(tmp.inf),apply(tmp.inf,1,sum)))

  tmp.inf <- sim.dodd.nomix[[i]]$all.pop
  tmp.inf <- tmp.inf[,2:ncol(tmp.inf)]
  num.inf.age.nomix.dodd$total.infections[((i-1)*(num.age+1)+1):(i*(num.age+1))] <- 
    round(c(sum(tmp.inf),apply(tmp.inf,1,sum)))

  ## par sweep
  tmp.inf <- sim.sweep.mix[[i]]$all.pop
  tmp.inf <- tmp.inf[,2:ncol(tmp.inf)]
  ### total.infections: (total  infections, age specific number of infections)
  num.inf.age.mix.sweep$total.infections[((i-1)*(num.age+1)+1):(i*(num.age+1))] <- round(c(sum(tmp.inf),apply(tmp.inf,1,sum)))

  tmp.inf <- sim.sweep.nomix[[i]]$all.pop
  tmp.inf <- tmp.inf[,2:ncol(tmp.inf)]
  num.inf.age.nomix.sweep$total.infections[((i-1)*(num.age+1)+1):(i*(num.age+1))] <- 
    round(c(sum(tmp.inf),apply(tmp.inf,1,sum)))
  
  # total number of infected individuals by high/low risk category
  ##################################################################
  ## Dodd Assumption
  tmp.inf.lo <- sim.dodd.mix[[i]]$low.risk.pop
  tmp.inf.lo <- tmp.inf.lo[,2:ncol(tmp.inf.lo)]
  tmp.inf.hi <- sim.dodd.mix[[i]]$high.risk.pop
  tmp.inf.hi <- tmp.inf.hi[,2:ncol(tmp.inf.hi)]
  num.inf.hi.low.dodd$total.infections[((i-1)*2+1):(i*2)] <- round(c(sum(tmp.inf.hi),sum(tmp.inf.lo)))
  
  ## Par sweep
  tmp.inf.lo <- sim.sweep.mix[[i]]$low.risk.pop
  tmp.inf.lo <- tmp.inf.lo[,2:ncol(tmp.inf.lo)]
  tmp.inf.hi <- sim.sweep.mix[[i]]$high.risk.pop
  tmp.inf.hi <- tmp.inf.hi[,2:ncol(tmp.inf.hi)]
  num.inf.hi.low.sweep$total.infections[((i-1)*2+1):(i*2)] <- round(c(sum(tmp.inf.hi),sum(tmp.inf.lo)))

  # total number of individuals with 0, 1,...,9+ infections by prevalence
  ##########################################################################
  ## Dodd assumption
  tmp.inf <- sim.dodd.mix[[i]]$all.pop
  ### compact together all individuals with 9+ infections
  tmp.inf <- cbind(tmp.inf[,1:9],apply(tmp.inf[,10:ncol(tmp.inf)],1,sum)) 
  num.inf.mix.dodd$count[((i-1)*10+1):(10*i)] <- round(apply(tmp.inf,2,sum))
 
  tmp.inf <- sim.dodd.nomix[[i]]$all.pop
  if(ncol(tmp.inf)<11){
    tmp <- as.data.frame(cbind(tmp.inf,matrix(rep(0,nrow(tmp.inf)*(11-ncol(tmp.inf))),nr=nrow(tmp.inf))))
    names(tmp)[(ncol(tmp.inf)+1):11] <- as.character((ncol(tmp.inf)+1):11) 
    tmp.inf <- tmp
  }
  # compact together all individuals with 9+ infections
  tmp.inf <- cbind(tmp.inf[,1:9],apply(tmp.inf[,10:ncol(tmp.inf)],1,sum)) 
  num.inf.nomix.dodd$count[((i-1)*10+1):(10*i)] <- round(apply(tmp.inf,2,sum))
  
  ## Par sweep
  tmp.inf <- sim.sweep.mix[[i]]$all.pop
  ### compact together all individuals with 9+ infections
  tmp.inf <- cbind(tmp.inf[,1:9],apply(tmp.inf[,10:ncol(tmp.inf)],1,sum)) 
  num.inf.mix.sweep$count[((i-1)*10+1):(10*i)] <- round(apply(tmp.inf,2,sum))
 
  tmp.inf <- sim.sweep.nomix[[i]]$all.pop
  if(ncol(tmp.inf)<11){
    tmp <- as.data.frame(cbind(tmp.inf,matrix(rep(0,nrow(tmp.inf)*(11-ncol(tmp.inf))),nr=nrow(tmp.inf))))
    names(tmp)[(ncol(tmp.inf)+1):11] <- as.character((ncol(tmp.inf)+1):11) 
    tmp.inf <- tmp
  }
  # compact together all individuals with 9+ infections
  tmp.inf <- cbind(tmp.inf[,1:9],apply(tmp.inf[,10:ncol(tmp.inf)],1,sum)) 
  num.inf.nomix.sweep$count[((i-1)*10+1):(10*i)] <- round(apply(tmp.inf,2,sum))

  # total number of individuals with 0,...9+ infections by prevalence, high/low
  ###############################################################################
  ## Dodd assumption
  tmp.inf <- sim.dodd.mix[[i]]$low.risk.pop
  # compact together all individuals with 9+ infections
  tmp.inf <- cbind(tmp.inf[,1:9],apply(tmp.inf[,10:ncol(tmp.inf)],1,sum)) 
  num.inf.mix.lo.dodd$count[((i-1)*10+1):(10*i)] <- round(apply(tmp.inf,2,sum))
  
  tmp.inf <- sim.dodd.mix[[i]]$high.risk.pop
  # compact together all individuals with 9+ infections
  tmp.inf <- cbind(tmp.inf[,1:9],apply(tmp.inf[,10:ncol(tmp.inf)],1,sum)) 
  num.inf.mix.hi.dodd$count[((i-1)*10+1):(10*i)] <- round(apply(tmp.inf,2,sum))
  
  ## Par sweep
  tmp.inf <- sim.sweep.mix[[i]]$low.risk.pop
  # compact together all individuals with 9+ infections
  tmp.inf <- cbind(tmp.inf[,1:9],apply(tmp.inf[,10:ncol(tmp.inf)],1,sum)) 
  num.inf.mix.lo.sweep$count[((i-1)*10+1):(10*i)] <- round(apply(tmp.inf,2,sum))
  
  tmp.inf <- sim.sweep.mix[[i]]$high.risk.pop
  # compact together all individuals with 9+ infections
  tmp.inf <- cbind(tmp.inf[,1:9],apply(tmp.inf[,10:ncol(tmp.inf)],1,sum)) 
  num.inf.mix.hi.sweep$count[((i-1)*10+1):(10*i)] <- round(apply(tmp.inf,2,sum))

  # prev of LTBI by age and TB prev
  ###########################################
  ## Dodd assumption
  tmp.inf <- sim.dodd.mix[[i]]$all.pop
  tmp.inf <- tmp.inf[,1] #numbers with no infections
  num.infect.age <- pop.age-tmp.inf
  age.prev.LTBI.mix.dodd[,i] <- round(num.infect.age/pop.age,2)

  tmp.inf <- sim.dodd.nomix[[i]]$all.pop
  tmp.inf <- tmp.inf[,1] #numbers with no infections
  num.infect.age <- pop.age-tmp.inf
  age.prev.LTBI.nomix.dodd[,i] <- round(num.infect.age/pop.age,2)

  ## Par sweep
  tmp.inf <- sim.sweep.mix[[i]]$all.pop
  tmp.inf <- tmp.inf[,1] #numbers with no infections
  num.infect.age <- pop.age-tmp.inf
  age.prev.LTBI.mix.sweep[,i] <- round(num.infect.age/pop.age,2)

  tmp.inf <- sim.sweep.nomix[[i]]$all.pop
  tmp.inf <- tmp.inf[,1] #numbers with no infections
  num.infect.age <- pop.age-tmp.inf
  age.prev.LTBI.nomix.sweep[,i] <- round(num.infect.age/pop.age,2)

}

max.inf.dodd <- 
  max(max(num.inf.age.nomix.dodd$total.infections),max(num.inf.age.mix.dodd$total.infections))
max.inf.sweep <- 
  max(max(num.inf.age.nomix.sweep$total.infections),max(num.inf.age.mix.sweep$total.infections))
max.inf <- max(max.inf.dodd,max.inf.sweep)
```


### Dodd, Homogeneous risk

This block of results uses the Dodd et al ARIs (determined from prevalence and using data from uninfected children) with the assumption of homogeneous risk of infection in the population.

```{r,echo=FALSE}
# plot the results for no mixing group
colorCount <- num.age
getPalette <- colorRampPalette(brewer.pal(9,"Paired"))

ggplot(data=num.inf.age.nomix.dodd[num.inf.age.nomix.dodd$age.group==0,],aes(x=prevalence,y=total.infections))+
  geom_bar(stat="identity")+ylim(0,max.inf+500)+
  labs(title="Total number of infected individuals",x="TB prevalence",y="Total number baseline infections")

ggplot(data=num.inf.age.nomix.dodd[num.inf.age.nomix.dodd$age.group!=0,],aes(x=as.factor(prevalence),y=total.infections,fill=as.factor(age.group)))+
  geom_bar(stat="identity")+ylim(0,max.inf+500)+
  labs(title="Total number of infected individuals by age and prevalence",x="TB prevalence",y="Total number baseline infections")+
  scale_fill_manual(name="Median age",values=getPalette(colorCount))+
  theme(legend.position = "bottom")+
  guides(fill=guide_legend(nrow=2))

ggplot(data=num.inf.nomix.dodd,aes(x=prevalence,y=count,fill=num.inf))+
  geom_bar(stat="identity")+
  labs(title="Distribution of the number of prior infections",x="TB prevalence",y="Total number of individuals")+
  scale_fill_brewer(name="Number of infections",palette="Paired")

```

We focus in on those with at least six prior infections in this plot.

```{r,echo=FALSE}
# just plot those with at least 6 infections
ggplot(data=num.inf.nomix.dodd[num.inf.nomix.dodd$num.inf>5,],aes(x=prevalence,y=count,fill=num.inf))+
  geom_bar(stat="identity")+
  labs(title="Distribution of the number of prior infections, at least 6 infections",x="TB prevalence",y="Total number of individuals")+
  scale_fill_brewer(name="Number of infections",palette="Paired")

```

Finally, we show the prevalence of LTBI by age and TB disease prevalence. This can be compared to the Dodd et al data.

```{r,echo=F}
row.names(age.prev.LTBI.nomix.dodd) <- c("0-4","5-9","10-14","15-19","20-24","25-29","30-34","35-39","40-44","45-49","50-54","55-59","60-64","65=-69","70-74","75-79","80-84","85+")

age.prev.LTBI.nomix.dodd %>% kbl(col.names = c("Prev100","Prev200","Prev300","Prev400","Prev500","Prev600","Prev700","Prev800","Prev900","Prev1000"),
                            caption="Prevalence of LTBI by age and TB prevalence") %>% kable_styling()
```

### Dodd, Heterogeneous risk

This section shows results using the Dodd ARIs (derived from prevalence data and using studies on unifected children). Here we assume that 20% of the population are in a high risk category and have an ARI that is 6 times higher than the remaining 80% of the population.

```{r, echo=F}
# plot the results for mixing group
ggplot(data=num.inf.age.mix.dodd[num.inf.age.mix.dodd$age.group==0,],aes(x=prevalence,y=total.infections))+
  geom_bar(stat="identity")+ylim(0,max.inf+500)+
  labs(title="Total number of infected individuals",x="TB prevalence",y="Total number baseline infections")

ggplot(data=num.inf.age.mix.dodd[num.inf.age.mix.dodd$age.group!=0,],aes(x=as.factor(prevalence),y=total.infections,fill=as.factor(age.group)))+
  geom_bar(stat="identity")+ylim(0,max.inf+500)+
  labs(title="Total number of infected individuals by age and prevalence",x="TB prevalence",y="Total number baseline infections")+
  scale_fill_manual(name="Median age",values=getPalette(colorCount))+
  theme(legend.position = "bottom")+
  guides(fill=guide_legend(nrow=2))

ggplot(data=num.inf.mix.dodd,aes(x=prevalence,y=count,fill=num.inf))+
  geom_bar(stat="identity")+
  labs(title="Distribution of the number of prior infections",x="TB prevalence",y="Total number of individuals")+
  scale_fill_brewer(name="Number of infections",palette="Paired")
```

We focus in on those with at least six prior infections in this plot.
```{r, echo=FALSE}
# just plot those with at least  infections
ggplot(data=num.inf.mix.dodd[num.inf.mix.dodd$num.inf>5,],aes(x=prevalence,y=count,fill=num.inf))+
  geom_bar(stat="identity")+
  labs(title="Distribution of the number of prior infections, at least 6 infections",x="TB prevalence",y="Total number of individuals")+
  scale_fill_brewer(name="Number of infections",palette="Paired")

```

We show the prevalence of LTBI by age and TB disease prevlance. This can be compared to the Dodd et al data.

```{r,echo=F}
row.names(age.prev.LTBI.mix.dodd) <- c("0-4","5-9","10-14","15-19","20-24","25-29","30-34","35-39","40-44","45-49","50-54","55-59","60-64","65=-69","70-74","75-79","80-84","85+")

age.prev.LTBI.mix.dodd %>% kbl(col.names = c("Prev100","Prev200","Prev300","Prev400","Prev500","Prev600","Prev700","Prev800","Prev900","Prev1000"),
                            caption="Prevalence of LTBI by age and TB prevalence") %>% kable_styling()
```

We also show the total number of infections and total number of times individuals have been infected by risk category.

```{r,echo=FALSE}
# add plot on for total number of infections by risk category
ggplot(data=num.inf.hi.low.dodd,aes(x=prevalence,y=total.infections,fill=as.factor(risk.cat)))+
  geom_bar(stat="identity")+
  labs(title="Total number of infected individuals by risk category",x="TB prevalence",y="Total number of infections")+
  scale_fill_brewer(name="Risk category",palette="Paired")

```
We can also consider the number of infections individuals have by prevalence and risk category.

```{r, echo=F}
ggplot(data=num.inf.mix.lo.dodd,aes(x=prevalence,y=count,fill=num.inf))+
  geom_bar(stat="identity")+
  labs(title="Distribution of the number of prior infections\nLow Risk",x="TB prevalence",y="Total number of individuals")+
  scale_fill_brewer(name="Number of infections",palette="Paired")

ggplot(data=num.inf.mix.hi.dodd,aes(x=prevalence,y=count,fill=num.inf))+
  geom_bar(stat="identity")+
  labs(title="Distribution of the number of prior infections\nHigh Risk",x="TB prevalence",y="Total number of individuals")+
  scale_fill_brewer(name="Number of infections",palette="Paired")

```

### Parameter sweep, Homogeneous risk

This block of results uses the parameter sweep determined ARI (determined assuming a two duration such that prevalence=2*incidence) with the assumption of homogeneous risk of infection in the population.

```{r,echo=FALSE}
# plot the results for no mixing group
colorCount <- num.age
getPalette <- colorRampPalette(brewer.pal(9,"Paired"))

ggplot(data=num.inf.age.nomix.sweep[num.inf.age.nomix.sweep$age.group==0,],aes(x=prevalence,y=total.infections))+
  geom_bar(stat="identity")+ylim(0,max.inf+500)+
  labs(title="Total number of infected individuals",x="TB prevalence",y="Total number baseline infections")

ggplot(data=num.inf.age.nomix.sweep[num.inf.age.nomix.sweep$age.group!=0,],aes(x=as.factor(prevalence),y=total.infections,fill=as.factor(age.group)))+
  geom_bar(stat="identity")+ylim(0,max.inf+500)+
  labs(title="Total number of infected individuals by age and prevalence",x="TB prevalence",y="Total number baseline infections")+
  scale_fill_manual(name="Median age",values=getPalette(colorCount))+
  theme(legend.position = "bottom")+
  guides(fill=guide_legend(nrow=2))

ggplot(data=num.inf.nomix.sweep,aes(x=prevalence,y=count,fill=num.inf))+
  geom_bar(stat="identity")+
  labs(title="Distribution of the number of prior infections",x="TB prevalence",y="Total number of individuals")+
  scale_fill_brewer(name="Number of infections",palette="Paired")

```

We focus in on those with at least six prior infections in this plot.

```{r,echo=FALSE}
# just plot those with at least 6 infections
ggplot(data=num.inf.nomix.sweep[num.inf.nomix.dodd$num.inf>5,],aes(x=prevalence,y=count,fill=num.inf))+
  geom_bar(stat="identity")+
  labs(title="Distribution of the number of prior infections, at least 6 infections",x="TB prevalence",y="Total number of individuals")+
  scale_fill_brewer(name="Number of infections",palette="Paired")

```

Finally, we show the prevalence of LTBI by age and TB disease prevalence. This can be compared to the Dodd et al data.

```{r,echo=F}
row.names(age.prev.LTBI.nomix.sweep) <- c("0-4","5-9","10-14","15-19","20-24","25-29","30-34","35-39","40-44","45-49","50-54","55-59","60-64","65=-69","70-74","75-79","80-84","85+")

age.prev.LTBI.nomix.sweep %>% kbl(col.names = c("Prev100","Prev200","Prev300","Prev400","Prev500","Prev600","Prev700","Prev800","Prev900","Prev1000"),
                            caption="Prevalence of LTBI by age and TB prevalence") %>% kable_styling()
```

### Parameter sweep, Heterogeneous risk


This block of results uses the parameter sweep determined ARI (determined assuming a two duration such that prevalence=2*incidence) with the assumption of homogeneous risk of infection in the population. Here we assume that 20% of the population are in a high risk category and have an ARI that is 6 times higher than the remaining 80% of the population.

```{r, echo=F}
# plot the results for mixing group
ggplot(data=num.inf.age.mix.sweep[num.inf.age.mix.sweep$age.group==0,],aes(x=prevalence,y=total.infections))+
  geom_bar(stat="identity")+ylim(0,max.inf+500)+
  labs(title="Total number of infected individuals",x="TB prevalence",y="Total number baseline infections")

ggplot(data=num.inf.age.mix.sweep[num.inf.age.mix.sweep$age.group!=0,],aes(x=as.factor(prevalence),y=total.infections,fill=as.factor(age.group)))+
  geom_bar(stat="identity")+ylim(0,max.inf+500)+
  labs(title="Total number of infected individuals by age and prevalence",x="TB prevalence",y="Total number baseline infections")+
  scale_fill_manual(name="Median age",values=getPalette(colorCount))+
  theme(legend.position = "bottom")+
  guides(fill=guide_legend(nrow=2))

ggplot(data=num.inf.mix.sweep,aes(x=prevalence,y=count,fill=num.inf))+
  geom_bar(stat="identity")+
  labs(title="Distribution of the number of prior infections",x="TB prevalence",y="Total number of individuals")+
  scale_fill_brewer(name="Number of infections",palette="Paired")
```

We focus in on those with at least six prior infections in this plot.
```{r, echo=FALSE}
# just plot those with at least  infections
ggplot(data=num.inf.mix.sweep[num.inf.mix.sweep$num.inf>5,],aes(x=prevalence,y=count,fill=num.inf))+
  geom_bar(stat="identity")+
  labs(title="Distribution of the number of prior infections, at least 6 infections",x="TB prevalence",y="Total number of individuals")+
  scale_fill_brewer(name="Number of infections",palette="Paired")

```

We show the prevalence of LTBI by age and TB disease prevlance. This can be compared to the Dodd et al data.

```{r,echo=F}
row.names(age.prev.LTBI.mix.sweep) <- c("0-4","5-9","10-14","15-19","20-24","25-29","30-34","35-39","40-44","45-49","50-54","55-59","60-64","65=-69","70-74","75-79","80-84","85+")

age.prev.LTBI.mix.sweep %>% kbl(col.names = c("Prev100","Prev200","Prev300","Prev400","Prev500","Prev600","Prev700","Prev800","Prev900","Prev1000"),
                            caption="Prevalence of LTBI by age and TB prevalence") %>% kable_styling()
```

We also show the total number of infections and total number of times individuals have been infected by risk category.

```{r,echo=FALSE}
# add plot on for total number of infections by risk category
ggplot(data=num.inf.hi.low.sweep,aes(x=prevalence,y=total.infections,fill=as.factor(risk.cat)))+
  geom_bar(stat="identity")+
  labs(title="Total number of infected individuals by risk category",x="TB prevalence",y="Total number of infections")+
  scale_fill_brewer(name="Risk category",palette="Paired")

```
We can also consider the number of infections individuals have by prevalence and risk category.

```{r, echo=F}
ggplot(data=num.inf.mix.lo.sweep,aes(x=prevalence,y=count,fill=num.inf))+
  geom_bar(stat="identity")+
  labs(title="Distribution of the number of prior infections\nLow Risk",x="TB prevalence",y="Total number of individuals")+
  scale_fill_brewer(name="Number of infections",palette="Paired")

ggplot(data=num.inf.mix.hi.sweep,aes(x=prevalence,y=count,fill=num.inf))+
  geom_bar(stat="identity")+
  labs(title="Distribution of the number of prior infections\nHigh Risk",x="TB prevalence",y="Total number of individuals")+
  scale_fill_brewer(name="Number of infections",palette="Paired")

```

## New cases generated over one year {.tabset}

We now generate cases over a one year period. All infected individuals with no reinfection in this year have a `r prog.old.inf` probability of progressing to active disease. Individuals with a first infection have a `r TB.prog[1]` probability of progressing. Those with a new re-infection follow the probability of progressing determined by Ackley. 

```{r,echo=F}

# simulate populations for each set of parameters and then get the basic summary statistics (not age specific)
## run the mixing scenarios ## 
##############################
res.print.mix.ackley.dodd <- res.print.mix.ackley.sweep <- matrix(0,nr=nrow(pars.mix),ncol=15)

for(i in 1:nrow(pars.mix)){
  # simulation new infections/cases
  #######################################
  ## Dodd assumption
  pars.tmp.dodd <- pars.mix[i,]
  
  if(is.na(pars.mix[i,1])){break}
  
  # simulate the initial population
  sim.dat.tmp <- sim.dodd.mix[[i]]
  
  # generate new cases
  max.inf <- ncol(sim.dat.tmp$all.pop)-1
  
  TB.prog.sim <- c(TB.prog[1],seq(TB.prog[2],TB.prog[2]+TB.prog[3]*(max.inf-1),TB.prog[3]))
  TB.prog.sim[TB.prog.sim>1] <- 1

  run.cases.dodd <- TB.sim(init.pop=sim.dat.tmp,TB.prog=TB.prog.sim,ari.high=pars.mix$ARI.high.risk[i]/100,ari.low=pars.mix$ARI.low.risk[i]/100,prog.old.inf)

  ## Par sweep
  pars.tmp.sweep <- pars.sweep[i,]
  
  if(is.na(pars.sweep[i,1])){break}
  
  # simulate the initial population
  sim.dat.tmp <- sim.sweep.mix[[i]]
  
  # generate new cases
  max.inf <- ncol(sim.dat.tmp$all.pop)-1
  
  TB.prog.sim <- c(TB.prog[1],seq(TB.prog[2],TB.prog[2]+TB.prog[3]*(max.inf-1),TB.prog[3]))
  TB.prog.sim[TB.prog.sim>1] <- 1

  run.cases.sweep <- 
    TB.sim(init.pop=sim.dat.tmp,TB.prog=TB.prog.sim,ari.high=pars.tmp.sweep$high.ARI..mix/100,ari.low=pars.tmp.sweep$low.ARI..mix/100,prog.old.inf)

  # get summary stats
  ##############################
  results.dodd <- sum.stats(run.cases.dodd,pars.tmp.dodd)
  results.sweep <- sum.stats(run.cases.sweep,pars.tmp.dodd) # pars here only uses prevalence, so use mix pars that are in correct format
  
  res.print.mix.ackley.dodd[i,] <-as.numeric(c(results.dodd$totals[,1],results.dodd$true.prev,results.dodd$incidence,results.dodd$duration,results.dodd$Rt.inf,results.dodd$Rt.dis))

  res.print.mix.ackley.sweep[i,] <- as.numeric(c(results.sweep$totals[,1],results.sweep$true.prev,results.sweep$incidence,results.sweep$duration,results.sweep$Rt.inf,results.sweep$Rt.dis))
  
}

colnames(res.print.mix.ackley.dodd) <- colnames(res.print.mix.ackley.sweep) <- c("low risk: ATB from first inf","low risk: ATB from reinfection", "low risk: total ATB", "high risk: ATB from first int","high risk: ATB from reinfection", "high risk: total ATB", "ATB from old infections",
                              "Total ATB","percent first inf","percent old inf", "corrected prevalence","incidence (per 100,000)","duration","Rt infections","Rt disease")

#################################
## run the no mixing scenarios ## 
#################################
res.print.nomix.ackley.dodd <- res.print.nomix.ackley.sweep <- matrix(0,nr=nrow(pars.nomix),ncol=9)
for(i in 1:nrow(pars.nomix)){
  ## Dodd assumption
  pars.tmp <- pars.nomix[i,]
  
  if(is.na(pars.nomix[i,1])){break}
  
  # simulate the initial population
  sim.dat.tmp <- sim.dodd.nomix[[i]]
  
  # generate new cases
  max.inf <- ncol(sim.dat.tmp$all.pop)-1
  
  TB.prog.sim <- c(TB.prog[1],seq(TB.prog[2],TB.prog[2]+TB.prog[3]*(max.inf-1),TB.prog[3]))
  TB.prog.sim[TB.prog.sim>1] <- 1

  run.cases.ackley.dodd <- TB.sim.nomix(init.pop=sim.dat.tmp,TB.prog=TB.prog.sim,ari=ari.overall.mix/100,prog.old.inf)

  ## Parameter sweep
  pars.tmp.sweep <- pars.sweep[i,]
  
  if(is.na(pars.nomix[i,1])){break}
  
  # simulate the initial population
  sim.dat.tmp <- sim.sweep.nomix[[i]]
  
  # generate new cases
  max.inf <- ncol(sim.dat.tmp$all.pop)-1
  
  TB.prog.sim <- c(TB.prog[1],seq(TB.prog[2],TB.prog[2]+TB.prog[3]*(max.inf-1),TB.prog[3]))
  TB.prog.sim[TB.prog.sim>1] <- 1

  run.cases.ackley.sweep <- TB.sim.nomix(init.pop=sim.dat.tmp,TB.prog=TB.prog.sim,ari=pars.tmp.sweep$overall.ARI..no.mix/100,prog.old.inf)

  ## get summary stats
  results.dodd <- sum.stats(run.cases.ackley.dodd,pars.tmp.dodd,mix=F)
  results.sweep <- sum.stats(run.cases.ackley.sweep,pars.tmp.dodd,mix=F)
  
  res.print.nomix.ackley.dodd[i,] <- as.numeric(c(results.dodd$totals[,1],results.dodd$true.prev,results.dodd$incidence,results.dodd$duration,results.dodd$Rt.inf,results.dodd$Rt.dis))

  res.print.nomix.ackley.sweep[i,] <- as.numeric(c(results.sweep$totals[,1],results.sweep$true.prev,results.sweep$incidence,results.sweep$duration,results.sweep$Rt.inf,results.sweep$Rt.dis))
  
}

colnames(res.print.nomix.ackley.dodd) <- colnames(res.print.nomix.ackley.sweep) <- c("ATB from old infections",
                              "Total ATB","percent first inf","percent old inf", "corrected prevalence","incidence (per 100,000)","duration","Rt infections","Rt disease")
```

We can calculate the following statistics (the first 6 only apply to the mixing scenarios):

- Cases from first infection, low risk
- Cases from reinfection, low risk
- Total cases, low risk
- Cases from first infection, high risk
- Cases from reinfection, high risk
- Total cases, high risk
- Cases from old infection (all in high risk)
- Total cases
- percent cases from first infection
- percent cases from old infection
- Corrected prevalence (=1.66*smear positive prevalence) 
- Incidence (per 100000)
- Implied duration (=prev/Inc)
- Reproductive number for infection (= #new infections/corrected prevalence)
- Reproductive number for disease (= #ATB from new infections/corrected prevalence)


### Dodd, homogeneous risk

```{R,echo=F}
  res.print.nomix.ackley.dodd %>% kbl() %>% kable_styling()
```

### Dodd, heterogeneous risk

```{r,echo=F}
  res.print.mix.ackley.dodd %>% kbl() %>% kable_styling()
```

### Parameter sweep, homogeneous risk
```{r,echo=F}
  res.print.nomix.ackley.sweep %>% kbl() %>% kable_styling()
```

### Parameter sweep, heterogeneous risk

```{r,echo=F}
  res.print.mix.ackley.sweep %>% kbl() %>% kable_styling()
```


### Plotting the results
We show the number of TB disease cases by prevalence and source (old infection, new primary infection, re-infection).

```{r,echo=FALSE}
print.nomix.dodd <- print.nomix.sweep <- print.mix.dodd <- print.mix.sweep <- data.frame(prevalence=rep(pars.mix[1:10,1],3),
                             source=rep(c("Old infection","New infection","Reinfection"),each=10),
                             count=rep(NA,30))

print.nomix.dodd$count <- c(res.print.nomix.ackley.dodd[,1],res.print.nomix.ackley.dodd[,3]*res.print.nomix.ackley.dodd[,2],
                              (1-(res.print.nomix.ackley.dodd[,3]+res.print.nomix.ackley.dodd[,4]))*res.print.nomix.ackley.dodd[,2])
print.nomix.sweep$count <- c(res.print.nomix.ackley.sweep[,1],res.print.nomix.ackley.sweep[,3]*res.print.nomix.ackley.sweep[,2],
                              (1-(res.print.nomix.ackley.sweep[,3]+res.print.nomix.ackley.sweep[,4]))*res.print.nomix.ackley.sweep[,2])

print.mix.dodd$count <- c(res.print.mix.ackley.dodd[,7],res.print.mix.ackley.dodd[,9]*res.print.mix.ackley.dodd[,8],
                              (1-(res.print.mix.ackley.dodd[,9]+res.print.mix.ackley.dodd[,10]))*res.print.mix.ackley.dodd[,8])
print.mix.sweep$count <- c(res.print.mix.ackley.sweep[,7],res.print.mix.ackley.sweep[,9]*res.print.mix.ackley.sweep[,8],
                              (1-(res.print.mix.ackley.sweep[,9]+res.print.mix.ackley.sweep[,10]))*res.print.mix.ackley.sweep[,8])

max.sweep <- max(sum(print.nomix.sweep$count[print.nomix.sweep$prevalence==1000]),sum(print.mix.sweep$count[print.mix.sweep$prevalence==1000]))
max.dodd <- max(sum(print.nomix.dodd$count[print.nomix.dodd$prevalence==1000]),sum(print.mix.dodd$count[print.mix.dodd$prevalence==1000]))

# create bar plots
ggplot(data=print.nomix.dodd,aes(x=prevalence,y=count,fill=source))+
  geom_bar(stat="identity")+ylim(c(0,max.dodd))+
  labs(title="Number of new cases\nNo mixing, Dodd ARI",x="TB prevalence",y="Number of cases")+
  scale_fill_brewer(name="Source of cases",palette="Paired")

ggplot(data=print.nomix.sweep,aes(x=prevalence,y=count,fill=source))+
  geom_bar(stat="identity")+ylim(c(0,max.sweep))+
  labs(title="Number of new cases\nNo mixing, Sweep ARI",x="TB prevalence",y="Number of cases")+
  scale_fill_brewer(name="Source of cases",palette="Paired")

ggplot(data=print.mix.dodd,aes(x=prevalence,y=count,fill=source))+
  geom_bar(stat="identity")+ylim(c(0,max.dodd))+
  labs(title="Number of new cases\nMixing, Dodd ARI",x="TB prevalence",y="Number of cases")+
  scale_fill_brewer(name="Source of cases",palette="Paired")

ggplot(data=print.mix.sweep,aes(x=prevalence,y=count,fill=source))+
  geom_bar(stat="identity")+ylim(c(0,max.sweep))+
  labs(title="Number of new cases\nMixing, Sweep ARI",x="TB prevalence",y="Number of cases")+
  scale_fill_brewer(name="Source of cases",palette="Paired")

```



