---
title: "ReMoTe simulation: Mixing Ackley"
author: ""
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r, include=FALSE}
source("./R/utils.R")
library(ggplot2)
library(reshape)
library(reshape2)
library(RColorBrewer)
library(ggpubr)
library(gridExtra)
library(grid)
library(scales)
library(dplyr)
library(data.table)
library(xlsx)
library(kableExtra)
```

## Background

We initialize a population of size one million using the ARI, and age structure that mimics South Africa. We then generate new case of disease over a one year period and explore the role of reinfection in generating new cases of disease. We simulate populations with prevalence ranging from 100/100k to 1000/100k.

We use the correction in our calculations that true prevalence should be 1.66 times reported to account for smear negative TB.

We explore varying two main parameters:
1. The risk of progression to disease as a function of the number of infections.
2. The proportion of the population that is at high risk of infection.

## Parameters used in this simulation
### Risk of progression to disease
Following is a depiction of the parameters used for the risk of progression to disease for those with a new infection as a function of the number of total infections.

```{r}
pars <- read.xlsx("G:/My Drive/Work/TB Projects/ReMoTe local/Parameters/Pars_2021_11_5.xlsx",sheetName = "Mixing and Ackley")
prevs <- pars[,1]
low.prev <- pars[1,3]
ari.goal <- c(0.9,1.27,1.63,2,2.36,2.72,3.1,3.46,3.83,4.2)/100
pop.age=c(99000, 99000,92000,83000,85000,90000, 91000,79000,63000,54000,44000,37000,30000, 23000,15000,10000,5000,1000)
median.age=c(2,7,12,17,22,27,32,37,42,47,52,57,62,67,72,77,82,87)
prog.old.inf=0.00075

max.inf <- 100
TB.prog<-cbind(pars$Progress1Infection[1],
              pars$Progression2Inf[1],
              pars$ProgressionStep[1])
TB.prog.sim <- c(TB.prog[1],rep(TB.prog[2],TB.prog[2]+TB.prog[3]*(max.inf-1),TB.prog[3]))
TB.prog.sim[TB.prog.sim>1] <- 1
```

Here, the risk of progression with one infection is `r TB.prog[1]` and the risk with two infections is `r TB.prog[2]`. After two infections the risk increases by `r TB.prog[3]` with each infection.

```{r}
plot.dat <- data.frame(num.infections=seq(1,max.inf+1),prog.rate=TB.prog.sim)

ggplot(plot.dat,aes(x=num.infections,y=prog.rate))+
  geom_line()+
  labs(x="Number of infections",y="Probability of pregression to disease")
```

### Proportion of the population at high risk

In this simulation, we assume that `r pars$PopLowRisk[1]` proportion of the population are at low risk of infection.

In summary the parameters for the simulation are:

```{r}
pars.sim <- pars
pars.sim %>% kbl(col.names=c("Prev(per100k)","Overall ARI","Population Low Risk","Prob progress with 1 infection",
                             "Probability with 2 infections","increase per addt'l infection")) %>% kable_styling()
```


## Parameter sweep to determine ARI

We perform a parameter sweep to determine the ARI that would create the observed ARI that Dodd used to correlate with prevalence. 

More specifically, Dodd et al determined that the ARIs that would correspond to the prevalences of 100/100k,..., 1000/100k are `r ari.goal`. These ARI were measured from uninfected children. Therefore, in our simulation, we seek to determine the true ARI that would produce this observed ARI among those who are uninfected at the start of our one year simulation. To do this for each prevalence assumed, we initialize the population using a broad range of ARIs and then generate cases. We count the number of individuals uninfected at the start of the simulation year and then the number of new cases among these individuals and calculate the ARI from these individuals and call this the observed ARI. We select the underlying ARI that produces an observed ARI that most closely matches the Dodd et al estimate. 

We note that the observed ARI for the scenario where individuals mix homogeneously should be the same as the Dodd estimate. This is because the ARI is uniformly applied to all uninfected at the start of the one year simulation and this becomes an unbiased estimator of the ARI.

When there is variation in the risk of exposure, as we have simulated in our heterogeneous mixing scenario, then the ARI is biased. If we assume that $p^{lo}$ and $p^{hi}$ of the high and low risk populations, respectively, are uninfected at the start of the simulation, then these are described as the probability of escaping infection for each year an individual is alive. For instance, if we have $N^{lo}_a$ individuals that are $a$ years old in the low risk group, then 

$p^{lo}=\sum_a^A N_a (1-\lambda^{lo})^a.$

This is similarly defined for $p^{hi}$. The observed ARI in any given year among the uninfected is

$\lambda^{obs}=\frac{\text{New Infections at time t+1}}{\text{Number uninfected at time t}}=\frac{\lambda^{lo}p^{lo}0.8+\lambda^{hi}p^{hi}0.2}{0.8p^{lo}+0.2p^{hi}}.$

Because $p^{lo] \ne p^{hi}$, this is not comparable to the ARI that generates the infections in the simulation. 

### ARIs estimated from parameter sweep
We perform a sweep through many candidate ARI values to determine the underlying ARI that would produce observed ARIs that are used by Dodd to link the ARI to prevalence. Following are the results.

```{r}
# get ARI that best matches for this par set
ARI.sweep <- runSweep(mix=T,low.prev=low.prev,TB.prog=as.numeric(pars[1,4:6]))

# output of runSweep() is: c("Prev(per100k)","Correct Prev","overall ARI","low risk ARI","high risk ARI","ARI achieved")  

ARI.sweep %>% kbl(col.names=c("Prev(per100k)","Correct Prev","overall ARI",
                              "low risk ARI","high risk ARI","ARI observed")) %>% kable_styling()


```


## Initializing the population 

We first determine the number of infections that individuals have at the start of the simulation. 

```{r, echo=F}
# simulate the baseline population
sim.data <- list()
for(i in 1:length(pars[,1])){
  
  sim.data[[i]] <- init.popn.mix(ari.high=ARI.sweep[i,5]/100,
                                ari.low=ARI.sweep[i,4]/100,
                                low.prev=pars[1,3])
}  
```

```{r, echo=FALSE}
# first show overall number of infections by prevalence
# need matrix with prevalence and number of infections, then have this by age group
num.prev <- length(unique(prevs)); num.age <-length(median.age)
# data frame for plotting with age,prev,total infections
num.inf.age <- data.frame(prevalence=rep(prevs,each=num.age+1),
                             age.group=rep(c(0,median.age),num.prev),
                             total.infections=rep(NA,num.prev*(num.age+1)))

# data frame with prevalence,number of infections (categorical), count with that number of infections/prev
num.inf.mix <- num.inf.mix.hi <- num.inf.mix.lo <-  
  data.frame(prevalence=rep(prevs,each=10),
                             num.inf=rep(c(as.character(c(0:8)),"9+"),num.prev),
                             count=rep(NA,num.prev*10))

# data frame for high/low and number of infections
num.inf.hi.low<- data.frame(prevalence=rep(prevs,each=2),
                             risk.cat=rep(c("High","Low"),10),
                             total.infections=rep(NA,num.prev*2))

# data frame prev of LTBI for each prevalenceTB x age combo
age.prev.LTBI <- matrix(0,nr=num.age,nc=10)

for(i in 1:10){
  
  # total number of infected individuals in each age group by prevalence
  #########################################################################
  tmp.inf <- sim.data[[i]]$all.pop
  tmp.inf <- tmp.inf[,2:ncol(tmp.inf)]
  ### total.infections: (total  infections, age specific number of infections)
  num.inf.age$total.infections[((i-1)*(num.age+1)+1):(i*(num.age+1))] <- round(c(sum(tmp.inf),apply(tmp.inf,1,sum)))

  # total number of infected individuals by high/low risk category
  ##################################################################
  tmp.inf.lo <- sim.data[[i]]$low.risk.pop
  tmp.inf.lo <- tmp.inf.lo[,2:ncol(tmp.inf.lo)]
  tmp.inf.hi <- sim.data[[i]]$high.risk.pop
  tmp.inf.hi <- tmp.inf.hi[,2:ncol(tmp.inf.hi)]
  num.inf.hi.low$total.infections[((i-1)*2+1):(i*2)] <- round(c(sum(tmp.inf.hi),sum(tmp.inf.lo)))
  
  # total number of individuals with 0, 1,...,9+ infections by prevalence
  ##########################################################################
  tmp.inf <- sim.data[[i]]$all.pop
  ### compact together all individuals with 9+ infections
  tmp.inf <- cbind(tmp.inf[,1:9],apply(tmp.inf[,10:ncol(tmp.inf)],1,sum)) 
  num.inf.mix$count[((i-1)*10+1):(10*i)] <- round(apply(tmp.inf,2,sum))
 
  # total number of individuals with 0,...9+ infections by prevalence, high/low
  ###############################################################################
  ## Dodd assumption
  tmp.inf <- sim.data[[i]]$low.risk.pop
  # compact together all individuals with 9+ infections
  tmp.inf <- cbind(tmp.inf[,1:9],apply(tmp.inf[,10:ncol(tmp.inf)],1,sum)) 
  num.inf.mix.lo$count[((i-1)*10+1):(10*i)] <- round(apply(tmp.inf,2,sum))
  
  tmp.inf <- sim.data[[i]]$high.risk.pop
  # compact together all individuals with 9+ infections
  tmp.inf <- cbind(tmp.inf[,1:9],apply(tmp.inf[,10:ncol(tmp.inf)],1,sum)) 
  num.inf.mix.hi$count[((i-1)*10+1):(10*i)] <- round(apply(tmp.inf,2,sum))
  
  # prev of LTBI by age and TB prev
  ###########################################
  ## Dodd assumption
  tmp.inf <- sim.data[[i]]$all.pop
  tmp.inf <- tmp.inf[,1] #numbers with no infections
  num.infect.age <- pop.age-tmp.inf
  age.prev.LTBI[,i] <- round(num.infect.age/pop.age,2)

}

max.inf <- max(num.inf.age$total.infections)
```


### Results of initialized population


```{r, echo=F}
# plot the results for mixing group
colorCount <- num.age
getPalette <- colorRampPalette(brewer.pal(9,"Paired"))

ggplot(data=num.inf.age[num.inf.age$age.group==0,],aes(x=prevalence,y=total.infections))+
  geom_bar(stat="identity")+ylim(0,max.inf+500)+
  labs(title="Total number of infected individuals",x="TB prevalence",y="Total number baseline infections")

ggplot(data=num.inf.age[num.inf.age$age.group!=0,],aes(x=as.factor(prevalence),y=total.infections,fill=as.factor(age.group)))+
  geom_bar(stat="identity")+ylim(0,max.inf+500)+
  labs(title="Total number of infected individuals by age and prevalence",x="TB prevalence",y="Total number baseline infections")+
  scale_fill_manual(name="Median age",values=getPalette(colorCount))+
  theme(legend.position = "bottom")+
  guides(fill=guide_legend(nrow=2))

ggplot(data=num.inf.mix,aes(x=prevalence,y=count,fill=num.inf))+
  geom_bar(stat="identity")+
  labs(title="Distribution of the number of prior infections",x="TB prevalence",y="Total number of individuals")+
  scale_fill_brewer(name="Number of infections",palette="Paired")
```

We focus in on those with at least six prior infections in this plot.
```{r, echo=FALSE}
# just plot those with at least  infections
ggplot(data=num.inf.mix[num.inf.mix$num.inf>5,],aes(x=prevalence,y=count,fill=num.inf))+
  geom_bar(stat="identity")+
  labs(title="Distribution of the number of prior infections, at least 6 infections",x="TB prevalence",y="Total number of individuals")+
  scale_fill_brewer(name="Number of infections",palette="Paired")

```

We show the prevalence of LTBI by age and TB disease prevalence. This can be compared to the Dodd et al data.

```{r,echo=F}
row.names(age.prev.LTBI) <- c("0-4","5-9","10-14","15-19","20-24","25-29","30-34","35-39","40-44","45-49","50-54","55-59","60-64","65-69","70-74","75-79","80-84","85+")

age.prev.LTBI %>% kbl(col.names = c("Prev100","Prev200","Prev300","Prev400","Prev500","Prev600","Prev700","Prev800","Prev900","Prev1000"),
                            caption="Prevalence of LTBI by age and TB prevalence") %>% kable_styling()
```

We also show the total number of infections and total number of times individuals have been infected by risk category.

```{r,echo=FALSE}
# add plot on for total number of infections by risk category
ggplot(data=num.inf.hi.low,aes(x=prevalence,y=total.infections,fill=as.factor(risk.cat)))+
  geom_bar(stat="identity")+
  labs(title="Total number of infected individuals by risk category",x="TB prevalence",y="Number of infected individuals")+
  scale_fill_brewer(name="Risk category",palette="Paired")

```
We can also consider the number of infections individuals have by prevalence and risk category.

```{r, echo=F}
ggplot(data=num.inf.mix.lo,aes(x=prevalence,y=count,fill=num.inf))+
  geom_bar(stat="identity")+
  labs(title="Distribution of the number of prior infections\nLow Risk",x="TB prevalence",y="Total number of individuals")+
  scale_fill_brewer(name="Number of infections",palette="Paired")

ggplot(data=num.inf.mix.hi,aes(x=prevalence,y=count,fill=num.inf))+
  geom_bar(stat="identity")+
  labs(title="Distribution of the number of prior infections\nHigh Risk",x="TB prevalence",y="Total number of individuals")+
  scale_fill_brewer(name="Number of infections",palette="Paired")

```


## New cases generated over one year

We now generate cases over a one year period. All infected individuals with no reinfection in this year have a `r prog.old.inf` probability of progressing to active disease. Individuals with a first infection have a `r TB.prog[1]` probability of progressing. Those with a new re-infection follow the probability of progressing determined by Ackley, which increases linearly from `r TB.prog[2]` by a step of `r TB.prog[3]` until it reaches 1. 

```{r,echo=F}

# simulate populations for each set of parameters and then get the basic summary statistics (not age specific)
## run the mixing scenarios ## 
##############################
res.print <- matrix(0,nr=length(prevs),ncol=16)
ari.calcs <- matrix(0,nr=length(prevs),nc=10) #store data to calculate observed ARI

for(i in 1:length(prevs)){
  # simulation new infections/cases
  #######################################
  # simulate the initial population
  sim.dat.tmp <- sim.data[[i]]
  
  # generate new cases
  max.inf <- ncol(sim.dat.tmp$all.pop)-1
  
  TB.prog.sim <- c(TB.prog[1],seq(TB.prog[2],TB.prog[2]+TB.prog[3]*(max.inf-1),TB.prog[3]))
  TB.prog.sim[TB.prog.sim>1] <- 1

  ari.hi  <- ARI.sweep[i,5]/100; ari.lo <- ARI.sweep[i,4]/100
  run.cases <- TB.sim(init.pop=sim.dat.tmp,TB.prog=TB.prog.sim,ari.high=ari.hi,ari.low=ari.lo,prog.old.inf)

  # get summary stats
  ##############################
  results.sum <- sum.stats(run.cases,prevs[i])

  res.print[i,] <-as.numeric(c(prevs[i],results.sum$totals[,1],results.sum$true.prev,results.sum$incidence,
                               results.sum$duration,results.sum$Rt.inf,results.sum$Rt.dis))


  # get num.uninfected, new infections, ARI
  # number uninfected, number new infections, ARI observed
  ##########################################################
  uninfected <- c(sum(run.cases$init.pop$all.pop[,1]),
                  sum(run.cases$init.pop$low.risk.pop[,1]),
                  sum(run.cases$init.pop$high.risk.pop[,1]))
  new.inf <- c(sum(run.cases$new.inf.pop.t1[,2]),
                 sum(run.cases$low.pop.t1[,2]),
               sum(run.cases$high.pop.t1[,2]))
  ari.calcs[i,] <- c(prevs[i],uninfected,new.inf,new.inf/uninfected)

}

colnames(res.print) <- c("Prevalence (/100k)","low risk: ATB from first inf","low risk: ATB from reinfection", "low risk: total ATB", "high risk: ATB from first int","high risk: ATB from reinfection", "high risk: total ATB", "ATB from old infections",
                              "Total ATB","percent first inf","percent old inf", "corrected prevalence","incidence (per 100,000)","duration","Rt infections","Rt disease")

colnames(ari.calcs) <- c("Prevalence","Overall: Uninfected","Low: Uninfected","High: Uninfected",
                                                           "Overall: New infections","Low: New infections","High: New Infections",
                                                           "Overall: Obs ARI","Low: Obs ARI","High: Obs ARI")

```

We can calculate the following statistics (the first 6 only apply to the mixing scenarios):

- Cases from first infection, low risk
- Cases from reinfection, low risk
- Total cases, low risk
- Cases from first infection, high risk
- Cases from reinfection, high risk
- Total cases, high risk
- Cases from old infection (all in high risk)
- Total cases
- percent cases from first infection
- percent cases from old infection
- Corrected prevalence (=1.66*smear positive prevalence) 
- Incidence (per 100000)
- Implied duration (=prev/Inc)
- Reproductive number for infection (= #new infections/corrected prevalence)
- Reproductive number for disease (= #ATB from new infections/corrected prevalence)

We also look at overall and by risk category, as appropriate:

- Uninfected individuals at start of simulation
- Number infected during one year
- Observed ARI

```{r,echo=F}
  res.print %>% kbl() %>% kable_styling()
```

We also calculate the observed ARI.

```{r}
   ari.calcs %>% kbl() %>% kable_styling()
```

### Plotting the results
We show the number of TB disease cases by prevalence and source (old infection, new primary infection, re-infection).

```{r,echo=FALSE}
print.mix <- data.frame(prevalence=rep(prevs,3),
                             source=rep(c("Old infection","New infection","Reinfection"),each=10),
                             count=rep(NA,30))

print.mix$count <- c(res.print[,8],res.print[,10]*res.print[,9],
                              (1-(res.print[,10]+res.print[,11]))*res.print[,9])

max.count <- max(sum(print.mix$count[print.mix$prevalence==1000]))

# create bar plots
ggplot(data=print.mix,aes(x=prevalence,y=count,fill=source))+
  geom_bar(stat="identity")+ylim(c(0,max.count))+
  labs(title="Number of new cases",x="TB prevalence",y="Number of cases")+
  scale_fill_brewer(name="Source of cases",palette="Paired")

# save data for plotting
save(num.inf.age,num.inf.mix,print.mix,file="./Output/Mix_ackley_plotData.RData")

```



