---
title: ""
author: "Me"
data: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

This run uses ARI values derived from parameter sweeps meant to preserve the relationship of prevalence being twice incidence. Here we assume that 80% of the population is in the low risk category regardless of prevalence.

This work reads in two functions which initialize and high and low risk population. The low risk population is subject to a lower risk of infection initially. The high risk population has a higher ARI. We calculate the number of infections these individuals have by age group and TB prevalence.

The second step is to generate new infections over a one year time period and then deterministically creates active cases of disease using prespecified progression rates that are a function of the number of infections an individual has.

```{r, echo = FALSE, include=FALSE}
setwd("G:/My Drive/Work/TB Projects/ReMoTe")
source("./R/utils.R")
library(ggplot2)
library(reshape)
library(reshape2)
library(RColorBrewer)
library(ggpubr)
library(gridExtra)
library(grid)
library(scales)
library(dplyr)
library(data.table)
library(xlsx)
library(kableExtra)
```

## Simulation Scenarios
We consider a set of prevalence values and associated ARI. We consider two scenarios about the population mixing: homogeneous mixing or a high and low risk exposure groups. 

### Mixing assumptions
Following are these parameter assumptions:

```{r, echo = FALSE}
pars.mix <- read.csv("G:/My Drive/Work/TB Projects/ReMoTe/Parameters/parSweepTest 2021-9-29.csv")

ari.overall.mix <- round(pars.mix$ARI.overall,2)

pars.print <- data.frame(cbind(pars.mix[,2],pars.mix$ARI.low,pars.mix$ARI.high,pars.mix$ARI.overall,rep(0.8,10),rep(0.2,10)))

pars.print %>% kbl(col.names=c("prevalence","mixing ARI low risk","mixing ARI high risk","overall ARI",
                               "prop. low risk","prop. high risk")) %>% kable_styling()

pop.age=c(99000, 99000,92000,83000,85000,90000, 91000,79000,63000,54000,44000,37000,30000, 23000,15000,10000,5000,1000)
median.age=c(2,7,12,17,22,27,32,37,42,47,52,57,62,67,72,77,82,87)
prog.old.inf=0.00075
```

### TB progression assumptions

We assume that the risk of TB progression varies with the number of prior infections. This assumes that there is protection, but that it wanes with an increasing number of infections (Ackley assumption). 

```{r, echo=F}
prog.1 <- 0.04
prog.2 <- 0.02
prog.step <- 0.01
max.inf <- 15

TB.prog <- c(prog.1,seq(prog.2,prog.2+prog.step*(max.inf-1),prog.step))
TB.prog[TB.prog>0.13] <- 0.13

plot.dat <- data.frame(num.infections=rep(seq(1,max.inf+1),1),assumption=rep(c("Ackley"),each=max.inf+1),prog.rate=TB.prog)

ggplot(data=plot.dat,aes(x=num.infections,y=prog.rate,group=assumption))+
  geom_line(aes(color=assumption))+
  geom_point(aes(color=assumption,shape=assumption))+
  labs(x="Number of infections",y="Probability of progression")

```


## Initializing the population

We first determine the number of infections that individuals have at the start of the simulation. 

```{r, echo=F}
# simulate the baseline population
sim.dat.mix <- list()
for(i in 1:length(pars.mix[,1])){
  
  if(is.na(pars.mix[i,1])){break}
  pars.tmp.mix <- pars.mix[i,]
  sim.dat.mix[[i]] <- init.popn.mix(ari.high=pars.tmp.mix$ARI.high/100,
                                ari.low=pars.tmp.mix$ARI.low/100,
                                low.prev=0.8)

}  
```

```{r, echo=FALSE}
# first show overall number of infections by prevalence
# need matrix with prevalence and number of infections, then have this by age group
num.prev <- length(pars.mix[1:10,2]); num.age <-length(median.age)
# data frame for plotting with age,prev,total infections
num.inf.age.mix <- data.frame(prevalence=rep(pars.mix[1:10,2],each=num.age+1),
                             age.group=rep(c(0,median.age),num.prev),
                             total.infections=rep(NA,num.prev*(num.age+1)))
# data frame with prevalence,number of infections (categorical), count with that number of infections/prev
num.inf.mix <- num.inf.mix.hi <- num.inf.mix.lo <- data.frame(prevalence=rep(pars.mix[1:10,1],each=10),
                             num.inf=rep(c(as.character(c(0:8)),"9+"),num.prev),
                             count=rep(NA,num.prev*10))
# data frame for high/low and number of infections
num.inf.hi.low <- data.frame(prevalence=rep(pars.mix[1:10,2],each=2),
                             risk.cat=rep(c("High","Low"),10),
                             total.infections=rep(NA,num.prev*2))
# data frame prev of LTBI for each prevalenceTB x age combo
age.prev.LTBI.mix <- matrix(0,nr=num.age,nc=10)

for(i in 1:10){
  
  # total number of infections in each age group by prevalence
  tmp.inf <- sim.dat.mix[[i]]$all.pop
  tmp.inf <- tmp.inf[,2:ncol(tmp.inf)]
  # total.infections: (total  infections, age specific number of infections)
  num.inf.age.mix$total.infections[((i-1)*(num.age+1)+1):(i*(num.age+1))] <- round(c(sum(tmp.inf),apply(tmp.inf,1,sum)))

  # total number of infections by high/low risk category
  tmp.inf.lo <- sim.dat.mix[[i]]$low.risk.pop
  tmp.inf.lo <- tmp.inf.lo[,2:ncol(tmp.inf.lo)]
  tmp.inf.hi <- sim.dat.mix[[i]]$high.risk.pop
  tmp.inf.hi <- tmp.inf.hi[,2:ncol(tmp.inf.hi)]
  num.inf.hi.low$total.infections[((i-1)*2+1):(i*2)] <- round(c(sum(tmp.inf.hi),sum(tmp.inf.lo)))
  
  # total number of individuals with 0, 1,...,9+ infections by prevalence
  tmp.inf <- sim.dat.mix[[i]]$all.pop
  # compact together all individuals with 9+ infections
  tmp.inf <- cbind(tmp.inf[,1:min(9,ncol(tmp.inf))],apply(tmp.inf[,10:ncol(tmp.inf)],1,sum)) 
  num.inf.mix$count[((i-1)*10+1):(10*i)] <- round(apply(tmp.inf,2,sum))
 
  # total number of individuals with 0,...9+ infections by prevalence, high/low
  tmp.inf <- sim.dat.mix[[i]]$low.risk.pop
  # compact together all individuals with 9+ infections
  tmp.inf <- cbind(tmp.inf[,1:9],apply(tmp.inf[,10:ncol(tmp.inf)],1,sum)) 
  num.inf.mix.lo$count[((i-1)*10+1):(10*i)] <- round(apply(tmp.inf,2,sum))
  
  tmp.inf <- sim.dat.mix[[i]]$high.risk.pop
  # compact together all individuals with 9+ infections
  tmp.inf <- cbind(tmp.inf[,1:9],apply(tmp.inf[,10:ncol(tmp.inf)],1,sum)) 
  num.inf.mix.hi$count[((i-1)*10+1):(10*i)] <- round(apply(tmp.inf,2,sum))
  
  # prev of LTBI by age and TB prev
  tmp.inf <- sim.dat.mix[[i]]$all.pop
  tmp.inf <- tmp.inf[,1] #numbers with no infections
  num.infect.age <- pop.age-tmp.inf
  age.prev.LTBI.mix[,i] <- round(num.infect.age/pop.age,2)
}

max.inf <- max(num.inf.age.mix$total.infections)
```


```{r, echo=F}
colorCount <- num.age
getPalette <- colorRampPalette(brewer.pal(9,"Paired"))

# plot the results for mixing group
ggplot(data=num.inf.age.mix[num.inf.age.mix$age.group==0,],aes(x=prevalence,y=total.infections))+
  geom_bar(stat="identity")+ylim(0,max.inf+500)+
  labs(title="Total number of infections",x="TB prevalence",y="Total number baseline infections")

ggplot(data=num.inf.age.mix[num.inf.age.mix$age.group!=0,],aes(x=as.factor(prevalence),y=total.infections,fill=as.factor(age.group)))+
  geom_bar(stat="identity")+ylim(0,max.inf+500)+
  labs(title="Total number of infections by age and prevalence",x="TB prevalence",y="Total number baseline infections")+
  scale_fill_manual(name="Median age",values=getPalette(colorCount))+
  theme(legend.position = "bottom")+
  guides(fill=guide_legend(nrow=2))

ggplot(data=num.inf.mix,aes(x=prevalence,y=count,fill=num.inf))+
  geom_bar(stat="identity")+
  labs(title="Distribution of the number of prior infections",x="TB prevalence",y="Total number of individuals")+
  scale_fill_brewer(name="Number of infections",palette="Paired")
```

We focus in on those with at least six prior infections in this plot.
```{r, echo=FALSE}
# just plot those with at least  infections
ggplot(data=num.inf.mix[num.inf.mix$num.inf>5,],aes(x=prevalence,y=count,fill=num.inf))+
  geom_bar(stat="identity")+
  labs(title="Distribution of the number of prior infections, at least 6 infections",x="TB prevalence",y="Total number of individuals")+
  scale_fill_brewer(name="Number of infections",palette="Paired")

```

We show the prevalence of LTBI by age and TB disease prevalance. This can be compared to the Dodd et al data.

```{r,echo=F}
row.names(age.prev.LTBI.mix) <- c("0-4","5-9","10-14","15-19","20-24","25-29","30-34","35-39","40-44","45-49","50-54","55-59","60-64","65=-69","70-74","75-79","80-84","85+")

age.prev.LTBI.mix %>% kbl(col.names = c("Prev100","Prev200","Prev300","Prev400","Prev500","Prev600","Prev700","Prev800","Prev900","Prev1000"),
                            caption="Prevalence of LTBI by age and TB prevalence") %>% kable_styling()
```

We also show the total number of infections and total number of times individuals have been infected by risk category.

```{r,echo=FALSE}
# add plot on for total number of infections by risk category
ggplot(data=num.inf.hi.low,aes(x=prevalence,y=total.infections,fill=as.factor(risk.cat)))+
  geom_bar(stat="identity")+
  labs(title="Total number of infections by risk category",x="TB prevalence",y="Total number of infections")+
  scale_fill_brewer(name="Risk category",palette="Paired")

```
We can also consider the number of infections individuals have by prevalence and risk category.

```{r, echo=F}
ggplot(data=num.inf.mix.lo,aes(x=prevalence,y=count,fill=num.inf))+
  geom_bar(stat="identity")+
  labs(title="Distribution of the number of prior infections\nLow Risk",x="TB prevalence",y="Total number of individuals")+
  scale_fill_brewer(name="Number of infections",palette="Paired")

ggplot(data=num.inf.mix.hi,aes(x=prevalence,y=count,fill=num.inf))+
  geom_bar(stat="identity")+
  labs(title="Distribution of the number of prior infections\nHigh Risk",x="TB prevalence",y="Total number of individuals")+
  scale_fill_brewer(name="Number of infections",palette="Paired")

```

## New cases generated over one year

We now generate cases over a one year period. All infected individuals with no reinfection in this year have a `r prog.old.inf` probability of progressing to active disease. Individuals with a first infection have a `r TB.prog[1]` probability of progressing.  

```{r,echo=F}

# simulate populations for each set of parameters and then get the basic summary statistics (not age specific)
## run the mixing scenarios ## 
##############################
res.print.mix.ackley <- matrix(0,nr=nrow(pars.mix),ncol=15)
for(i in 1:nrow(pars.mix)){
  pars.tmp <- pars.mix[i,]
  
  if(is.na(pars.mix[i,1])){break}
  
  # simulate the initial population
  sim.dat.tmp <- sim.dat.mix[[i]]
  
  # generate new cases
  max.inf <- ncol(sim.dat.tmp$all.pop)-1
  
  TB.prog <- c(prog.1,seq(prog.2,prog.2+prog.step*(max.inf-1),prog.step))
  TB.prog[TB.prog>0.13] <- 0.13

  run.cases.ackley <- TB.sim(init.pop=sim.dat.tmp,TB.prog=TB.prog,ari.high=pars.mix$ARI.high[i]/100,ari.low=pars.mix$ARI.low[i]/100,prog.old.inf)

  # get summary stats
  results.sum.ackley <- sum.stats(run.cases.ackley,c(pars.tmp$X.1,pars.tmp$ARI.low,pars.tmp$ARI.high,0.80,0.20))

  res.print.mix.ackley[i,] <- c(results.sum.ackley$totals[,1],results.sum.ackley$true.prev,results.sum.ackley$incidence,results.sum.ackley$duration,results.sum.ackley$Rt.inf,results.sum.ackley$Rt.dis)

}

colnames(res.print.mix.ackley) <- c("low risk: ATB from first inf","low risk: ATB from reinfection", "low risk: total ATB", "high risk: ATB from first int","high risk: ATB from reinfection", "high risk: total ATB", "ATB from old infections",
                              "Total ATB","percent first inf","percent old inf", "corrected prevalence","incidence (per 100,000)","duration","Rt infections","Rt disease")


```

We can calculate the following statistics (the first 6 only apply to the mixing scenarios):

- Cases from first infection, low risk
- Cases from reinfection, low risk
- Total cases, low risk
- Cases from first infection, high risk
- Cases from reinfection, high risk
- Total cases, high risk
- Cases from old infection (all in high risk)
- Total cases
- percent cases from first infection
- percent cases from old infection
- Corrected prevalence (=1.66*smear positive prevalence) 
- Incidence (per 100000)
- Implied duration (=prev/Inc)
- Reproductive number for infection (= #new infections/corrected prevalence)
- Reproductive number for disease (= #ATB from new infections/corrected prevalence)


### Mixing, Ackley

```{R,echo=F}
  res.print.mix.ackley %>% kbl() %>% kable_styling()
```


### Plotting the results
We show the number of TB disease cases by prevalence and source (old infection, new primary infection, re-infection).

```{r,echo=FALSE}
print.mix.ackley <- data.frame(prevalence=rep(pars.mix[1:10,1],3),
                             source=rep(c("Old infection","New infection","Reinfection"),each=10),
                             count=rep(NA,30))

print.mix.ackley$count <- c(res.print.mix.ackley[,7],res.print.mix.ackley[,9]*res.print.mix.ackley[,8],
                              (1-(res.print.mix.ackley[,9]+res.print.mix.ackley[,10]))*res.print.mix.ackley[,8])

# create bar plots
ggplot(data=print.mix.ackley,aes(x=prevalence,y=count,fill=source))+
  geom_bar(stat="identity")+
  labs(title="Number of new cases\nMixing, Ackley",x="TB prevalence",y="Number of cases")+
  scale_fill_brewer(name="Source of cases",palette="Paired")

```



