---
title: "ReMoTe Sensitivity Analyses"
author: "Laura White"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Sensitivity analyses {.tabset}
Here we run the ReMoTe model. We first vary the proportion of the population at high risk of TB exposure and the probability of TB disease progression. In the latter, we consider one scenario with constant protection from progression (non-Ackley) and four scenarios with protection where we vary either the probability of progression with one infection or the rate that protection wanes with subsequent infection. Finally we consider the scenario where the difference between the ARI in the high and low risk groups varies between 4 and 8 (the main results have this set to 6).

```{r, include=F}
source('./utils.R')

prevs <- seq(100,1000,100)

# TB.prog scenarios
TB.prog1 <- c(0.05,0.02,0.01)
TB.prog2 <- c(0.04,0.02,0.01)
TB.prog3 <- c(0.06,0.02,0.01)
TB.prog4 <- c(0.05,0.02,0.005)
TB.prog5 <- c(0.05,0.02,0.0)
TB.prog.all <- rbind(TB.prog1,TB.prog2,TB.prog3,TB.prog4,TB.prog5)

# low.prev scenarios
low.prev.all <- seq(0.6,0.95,0.05)

# ARI ratio
ari.rat <- seq(4,9,1)

```


### Results: changing proportion at low risk

We vary the proportion at low risk of infection between 0.7 0.8, and 0.9. Following are results demonstrating the change in ARI and incidence for the 10 prevelance scenarios. These assume that TB progression rates are 0.05 for the first infection, drop to 0.02 for the second and increase by 0.01 for each subsequent infection.

```{r}
# low prevalence simulations-vary proportion at low risk of exposure #
######################################################################
low.prev.res <- list()
for(i in 1:length(low.prev.all)){
  low.prev.res[[i]] <- runAllSim(TB.prog=TB.prog1,low.prev=low.prev.all[i],mix=T)
}

# TB prog simulations-vary risk of disease progression #
########################################################
TB.prog.res <- list()
for(i in 1:length(TB.prog.all[,1])){
  TB.prog.res[[i]] <- runAllSim(TB.prog=TB.prog.all[i,],low.prev=0.8,mix=T)
}

# CREATE PLOTS OF RESULTS #
###########################
# change in low prev and incidence
######################################
incidences <- matrix(0,nr=length(low.prev.all),ncol=length(prevs))
aris <- matrix(0,nr=length(low.prev.all),ncol=length(prevs))
for(i in 1:length(low.prev.all)){
  incidences[i,] <- matrix(unlist(low.prev.res[[i]][[1]]),ncol=10)[1,]
  aris[i,] <- low.prev.res[[i]][[2]]
}

plot.dat <- data.frame(low.prev=rep(low.prev.all,10),Prevalence=rep(prevs,each=length(low.prev.all)),
                       ari=c(aris),incidence=c(incidences))

# incidence versus risk plot
ggplot(plot.dat,aes(x=Prevalence,y=incidence,color=low.prev))+geom_point()+
  labs(x="Prevalence",y="Observed incidence")+scale_color_continuous(name="Proportion\n Low Risk")


# ari versus risk plot
ggplot(plot.dat,aes(x=Prevalence,y=ari,color=low.prev))+geom_point()+
  labs(x="Prevalence",y="Actual ARI")+scale_color_continuous(name="Proportion\n Low Risk")

```

### Results: changing the risk of TB disease assumption

```{r}
## Change in risk of TB disease assumptions 
#############################################
incidences <- matrix(0,nr=length(TB.prog.all[,1]),ncol=length(prevs))
aris <- matrix(0,nr=length(TB.prog.all[,1]),ncol=length(prevs))
for(i in 1:length(TB.prog.all[,1])){
  incidences[i,] <- matrix(unlist(TB.prog.res[[i]][[1]]),ncol=10)[1,]
  aris[i,] <- TB.prog.res[[i]][[2]]
}

plot.dat.2 <- data.frame(Prevalence=rep(prevs,each=5),Ackley=rep(c("Yes","Yes","Yes","Yes","No"),10),
                         fast=rep(c(0.01,0.01,0.01,0.005,NA),10),Prog1=rep(c(0.05,0.04,0.06,0.05,0.05),10),
                         ari=c(aris),incidence=c(incidences))

# plot of prevalence versus incidence by ackley/no ackley
############################################################
plot.dat.3a <- subset(plot.dat.2,plot.dat.2$fast==0.01)
plot.dat.3b <- subset(plot.dat.2,is.na(plot.dat.2$fast))
plot.dat.3 <- data.frame(rbind(plot.dat.3a,plot.dat.3b))
ggplot(plot.dat.3,aes(x=Prevalence,y=incidence,shape=as.factor(Ackley),color=as.factor(Prog1)))+
  geom_point()+
  labs(y="Incidence",color="TB progression\n with one infection",shape="Waning protection")

plot.dat.4 <- subset(plot.dat.2,plot.dat.2$Prog1==0.05)
ggplot(plot.dat.4,aes(x=Prevalence,y=incidence,shape=as.factor(Ackley),color=as.factor(fast)))+
  geom_point()+
  labs(y="Incidence",color="Progression rate",shape="Waning protection")

```

### Results: changing ratio between high and low risk ARI

We vary the ratio between ARIs for the high and low risk populations from 4 to 8. In the main analyses, this is assumed to be 6. For these scenarios, we maintain 80% of the population in the low risk group.

```{r}
# vary ARI.rat #
################
ARI.rat.res <- list()
for(i in 1:length(ari.rat)){
  ARI.rat.res[[i]] <- runAllSim(ARI.rat=ari.rat[i],mix=T)
}

# CREATE PLOTS OF RESULTS #
###########################
# change in low prev and incidence
######################################
incidences <- matrix(0,nr=length(ari.rat),ncol=length(prevs))
aris <- matrix(0,nr=length(ari.rat),ncol=length(prevs))
for(i in 1:length(ari.rat)){
  incidences[i,] <- matrix(unlist(ARI.rat.res[[i]][[1]]),ncol=10)[1,]
  aris[i,] <- ARI.rat.res[[i]][[2]]
}

plot.dat <- data.frame(ari.rat=rep(ari.rat,10),Prevalence=rep(prevs,each=length(ari.rat)),
                       ari=c(aris),incidence=c(incidences))

# incidence versus risk plot
ggplot(plot.dat,aes(x=Prevalence,y=incidence,color=ari.rat))+geom_point()+
  labs(x="Prevalence",y="Observed incidence")+scale_color_continuous(name="ARI ratio")

# ari versus risk plot
ggplot(plot.dat,aes(x=Prevalence,y=ari,color=ari.rat))+geom_point()+
  labs(x="Prevalence",y="Actual ARI")+scale_color_continuous(name="ARI ratio")
```